Towards a Complete Model of the Negative Selection Process in Humans

### 2023-06-23 manuscript/manuscript.tex
ms updates

Our work is motivated by a well-known limitation of classic background
selection theory: it is only accurate for strongly deleterious mutations
\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}. Given
statistical methods that fit linked selection processes to genome-wide
diversity data rely on the classic BGS model to estimate the distribution of
fitness effects and rate of deleterious mutations, it is important to
characterize and remedy these shortcomings. However, extending background
selection theory to model weak negative selection is challenging, and it is
worthwhile to develop some intuition for why this is the case, how the classic
background selection works, and how the model of \textcite{Santiago2016-mu}
solves these problems and can be extended to genome-wide inference.

Both the classic background selection and SC16 models imagine a negative
selection process where a continual influx of deleterious mutations flow into
the population at a rate of $\mu$ per basepair per generation in a conserved
region of $L$ basepairs, such that the region-wide per generation mutation rate
is $U = \mu L$. Each mutation imposes an additive selective cost of $s$ in
heterozygotes and $2s$ in homozygotes. The classic background selection model
considers when mutations are sufficiently deleterious that they are destined to
extinction
\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1994-oh}. In
this case, the genealogy is well-approximated by a neutral coalescence process
with a rescaled effective population size of $N_0 = N \exp(-\nicefrac{U}{s})$.
This approximation works reasonably well for two reasons (though see
\cite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, when selection
is strong, the number of deleterious mutations per haplotype has a stationary
Poisson distribution with an average equal to the equilibrium under the
deterministic mutation-selection model \parencite{Haldane1927-ga}. Since
lineages carrying any strongly deleterious mutations cannot contribute to the
genealogy, coalescence events only occur in the fraction
$\exp(-\nicefrac{U}{s})$ of the population free of mutations. Second, the time
it takes for a present-day lineage to trace their ancestry back to a
mutation-free ancestor (the ``delay phase") is negligible compared to the
coalescence timescale of $N_0$ among these ancestors (the ``coalescence
phase"), and can be ignored (\cite{Durrett2008-ql}, p. 213;
\cite{Good2014-yz}).

When mutations are only weakly deleterious, their allele frequencies are
strongly influenced by stochastic perturbations and their frequency
trajectories are no longer well-approximated by deterministic models. When
genetic drift is the only source of randomness in allele frequency change, the
relevant scale of stochastic perturbations is determined by the drift-effective
population size $N_e$ (e.g. due to only non-selective process). The insight of
nearly neutral theory (XXX) was that the qualitative behavior of selected sites
is determined by the compound parameter $2N_e s$. When $2N_e s \gg 1$, the
stochastic component of allele frequency change is minuscule relative to
selection and can be ignored; this is why the factor that rescales the
drift-effective population size under the BGS model does not depend on $N_e$.
However, when $2N_e s \le 1$, the stochastic perturbations are on a scale equal
to or greater than the changes due to selection. In this regime, weakly
deleterious alleles can drift up to intermediate frequencies before their
eventual loss or fixation. When this occurs, classic BGS theory breaks down for
several reasons. First, the ``delay phase" is no longer negligible as lineages
carrying weakly deleterious mutations can persist on coalescent timescales.
This distorts genealogies away those expected under neutrality
\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. Second, the
distribution of the number of deleterious mutations (and its corresponding
fitness distribution) are no longer stationary and become a traveling wave
\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
population fitness (since beneficial mutations are ignored). Consequently,
individuals in the least-loaded class are stochastically lost, fixing those
mutations with a click of ``Muller's ratchet" \parencite{Muller1964-ki}.
Determining both the rate of Muller's ratchet
\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz} and the shape of
genealogies under weak selection have been stubborn open problems in
evolutionary genetics.

To further complicate matters, both strong and weakly deleterious mutations
create \emph{heritable} fitness variation. In turn, the presence of heritable
fitness variation generates another source of randomness in allele frequency
change known as genetic \emph{draft} \parencite{Neher2013-dz}. An allele
experiences genetic draft when it becomes randomly associated with a fitness
background, which stochastically perturbs its trajectory across generations.
Like genetic drift, perturbations due to draft are directionless but overall
increase the variance in allele frequency change. Unlike drift, stochastic
perturbations caused by draft are autocorrelated across generations for as long
an allele does not recombine or segregate off its fitness background and the
selective environment does not change
\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When selection is
strong, draft can generate multiple-merger coalescences that create
discontinuities in allele frequency trajectories, making diffusion
approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
However, when selection is weak, draft can be approximated as an increase in
the rate of drift (or equivalently, a reduction in effective population size).

A key insight from Robertson (\citeyear{Robertson1961-ho}) is that in the
long-run draft acts like an increase in the variance in offspring number.
\textcite{Wright1938-tv} showed that additional non-heritable variance in
offspring number (e.g. due to mating strategy) rescales the population size
according to,

@@

While using Equation \eqref{eq:Va} in Equation \eqref{eq:main_Ne} leads to a
prediction for the draft-effective population size $N_d$, closed-form
expressions for the rate of the ratchet $R$ have generally been hard to find
\parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. The key insight of
\textcite{Santiago2016-mu} is that the ratchet rate under draft is determined
by the probability of fixation $p_f(N_d, s)$
\parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
population size, i.e. $R = N U p_f(N_d, s)$. Given this equation for the
ratchet and Equation \eqref{eq:polygenic_Ne} for $N_d$ under draft, we have a
system of two non-linear equations that can be solved numerically for $N_d$ and
$R$ for each segment,

@@

where $V_a = (U-R)s$ as in Equation \eqref{eq:Va}.

@@

Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
haploid-model for a single region under equilibrium negative selection, we
first sought to confirm it could accurately predict the reduction in realistic
forward-time simulations of diploid populations with recombination. Using SLiM
\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
basepairs under varying levels of mutation and selection. We find a close
correspondence between the observed and predicted reduction in effective
population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
parameters including weak selection (Figure \ref{fig:figure-1}A), in contrast
to classic BGS theory. Furthermore, to investigate whether this accuracy was
caused by the model accurately predicting the equilibrium fitness variance and
ratchet rate, we also measured these throughout the simulation. Again, we find
the diploid SC16 theory accurately predicts both the ratchet rate (Figure
1\ref{fig:figure-1}B) and the genic fitness variance (Figure
1\ref{fig:figure-1}C).

Additionally, these figures provide further intuition about the underlying
negative selection process. When selection is strong the ratchet rate is zero,
as there is no probability that strongly deleterious can fix. However, around
$2N_e s \approx 1$, the ratchet begins clicking as $p_F > 0$. When this occurs,
fitness variation is reduced from the strong-selection equilibrium by the
amount of population fitness lost per generation due to the ratchet.
Ultimately, the ratchet rate reaches the neutral mutation rate, as $V_a = 0$
and $N_d = N$ (e.g. no draft, pure drift). Likewise, the additive genic fitness
variance matches the level predicted by BGS ($Us$), but begins to diverge when
the ratchet begins clicking. This occurs when the level of variation diverges
from the deterministic mutation-selection equilibrium $V_a = Us$ due to the
stochastic clicking of the ratchet at rate $R$.

@@

Given the accuracy of the SC16 model in predicting the reduction factor $B$ and
ratchet rate for a single segment under general equilibrium mutation negative
selection processes, we extended their model so that it could be applied to
whole-genome data. Our software method \texttt{bgspy} calculates the
equilibrium additive genic fitness variance ($V_a$) and ratchet rate ($R$)
across grids of mutation rates and selection coefficients for each
pre-specified segment in the genome that may be under negative selection (e.g.
coding sequences or UTRs). Then, our method computes a reduction map $B(x)$ for
each position $x$ in the genome by combining the pre-computed $V_a$ for every
segment and a user-supplied recombination map (see Methods XXX). To distinguish
between McVicker's B maps based on classic BGS theory, we call our reduction
map $B(x)$ a B' map.

To assess our predicted B' reduction maps, we simulated negative selection for
fixed mutation rates and selection coefficient on human chromosome 10 using
realistic recombination and conserved feature maps. Averaging over $n=100$
replicates, we estimated the simulation expected reduction map and compared
this to the predicted B and B' maps for the same parameters. We find that our
B' maps and the classic BGS theory B maps closely match simulations when
selection is strong (top row of Figure \ref{fig:figure-1}E). This provides the
first realistic forward-time chromosome-scale simulation confirmation of
classic BGS theory. However, we find slight discrepancies in regions with low
recombination (Figure \ref{fig:figure-1}E). Second, we find our theory is
vastly more accurate than the classic B maps when selection is very weak ($2N_e
s \ll 1$; bottom row of Figure \ref{fig:figure-1}E), which is expected since
BGS theory does not work in this domain. Across all mutation and selection
parameters simulated, the relative error of the classic B maps is 14.6\%
whereas the relative error in the new B' maps is 5\%. Additionally, we find
that the mean squared error between simulations and B' maps is close to the
theoretic lower bound set by the coalescence variance \parencite[see Methods
XXX]{Tajima1983-gu}.

Nearly all of the error between the new B' maps and chromosome-wide simulations
occurs around the drift-barrier domain of $2Ns \approx 1$ (Figure
\ref{fig:figure-1}D). The higher error in this domain occurs across all
mutation rates. We hypothesized that the higher error in this domain may be
because when our method numerically solves Equations \eqref{eq:main_eqns} for
each segment, it does not consider the experienced reduction due to selection
at other linked segments. In particular, we use a fixed drift-effective
$N=1000$ corresponding to the number of diploids in the simulations rather than
$B(x) N$ at position $x$. To test this, we implemented a locally-rescaled
version of the B' maps, which numerically solves Equations \eqref{eq:main_eqns}
for each segment at position $x$ taking into account the reduction experienced
due to selection at other segments by setting $B'(x)N$ as the population size
(see Method XXX). We find the locally-rescaled B' maps reduce the relative
error from 5\% to 0.4\% and mean squared error (Figure \ref{fig:figure-1}D,
dashed colored lines), but does not entirely eliminate the error in the $2Ns
\approx 1$ domain.

@@

Overall, while we find evidence that local-recalling can reduce error rate in
the $2Ns \approx 1$ domain,

@@

\begin{align}
    \label{eq:sub}
    r_f = \mu \int_0^1 p_f(s) ds
\end{align}

@@

\subsection*{Segment Simulations}

@@

\subsection*{Solving the B' Equations for each Segment}

In our implementation of the theory of \textcite{Santiago2016-mu}, we calculate
the local effective population size $N_e(x) = b(x) N$ within a segment $x$ by
solving a system of two non-linear equations,

\begin{align}
  \label{eq:}
  {N}_{e}(x) &= N \exp \left( -\frac{V(x)}{2} Q^2(x) \right) & \text{\emph{effective population equation}} \\
  {R}(x) &= \frac{4N_e(x)U s}{\exp(4 N_e(x) s)-1}  & \text{\emph{substitution equation}} 
\end{align}

\begin{align}
  %V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
  V(x) &= (U  - 2 {R}(x)) s & \text{\emph{fitness variance equation}} \\
  Q^2(x) &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
  Z &= 1 - \frac{Us}{U-{R(x)}} & \text{\emph{}} \text{\emph{variance decay rate}}
\end{align}

and $U = \mu L$ and $M = r_\text{BP} L$ are the total mutation and
recombination rates for a segment $L$ basepairs long. Each mutation has a
selection coefficient of $s$, for which we assume additive effects. Then, the
local reduction factor is $b(x) = \nicefrac{N_e(x)}{N}$, and the per-basepair
per-generation substitution rate is $\lambda_D = \nicefrac{R(x)}{L}$. Note that
in the variance equation, the long-run variance is the balance between mutation
and twice the loss due to substitution, $(U-2R)s$, since for selective balance
$V(x) = 0$ for $s>0$, the per-basepair substitution rate must be half the mutation rate,
since each substitution removes twice the variance introduced by a mutation in
a diploid species.

In our software \texttt{bgspy}, these two equations are solved for a grid of
mutation rates and selection coefficients for each specified segment from a
feature type (e.g. CDS, UTRs, etc.). This results in estimates of $N_e(x)$ and
$R(x)$ per segment. In the first-round B' calculation, we use a single
pre-specified drift-effective population size, $N$. The B' maps created here at
this step reflect selection only in the segment, ignoring the impact of
selection at other segments. By contrast, the locally-rescaled B' maps solve
these equations, assuming the drift-effective population size is $B(x)N$. These
reflect the segment's dynamics if they were effected by selection at other
segments under selection in the genome. Consequently, we can compare and
contrast models with local-recalling to test whether we achieve a better fit if
we account selective interference from other segments.

\subsection*{The B' Maps}

The reduction at a focal site $x$ can be calculated as the product of all
segments' contributing reduction factors,

\begin{align}
  B(v) = \prod_x b(x,v)
\end{align}

@@

\section*{Appendix}

\subsection*{Overview of Quantitative Genetic Models of Effective Population Size}

Here, we step through a quick derivation of \textcite{Santiago1995-hx}.
Throughout, we assume random mating, hermaphroditic individuals, and a constant
population size. The change in a neutral allele's frequency in one generation
can be partitioned into the three sources of stochasticity: the random
associations with fitness backgrounds (i.e. \emph{draft}), the non-heritable
randomness in family size, and the Mendelian noise from heterozygotes
segregating. If we let $x_{0,i} \in \{0, \nicefrac{1}{2}, 1\}$ be the frequency
of neutral alleles individual $i$ in generation 0 carries, we can partition the
random neutral allele frequency of the population ($x_t$ without the individual
index) into the underlying stochastic causes,

\begin{align}
  x_1 = \frac{1}{2N} \sum_{i=1}^N \left( x_{0,i}k_{0,i} + \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \right)
\end{align}

where $k_{0,i}$ is the number of surviving gametes parent $i$ passes on, and
$\delta_{0,i,j}$ is a random term that encapsulates the noise due to Mendelian
segregation of heterozygotes. If parent $i$ is a homozygote ($x_{0,i} \in \{0,
1\}$), then $\delta_{0,i,j} = 0$, whereas if $x_{0,i} = \nicefrac{1}{2}$ then
$\delta_{0,i,j} = \pm \nicefrac{1}{2}$ with equal probability. This is because
a heterozygous parent will transmit half a neutral allele in expectation, but
each round of Mendelian segregation must pass on either 0 or 1 alleles, which
the random $\pm \nicefrac{1}{2}$ term imposes. The factor of $\nicefrac{1}{2}$
is due to the fact that we're summing over $N$ diploids, but considering the
number of gametes they transmit. Since each diploid parent must have two
offspring to maintain a constant population size, $\nicefrac{1}{N} \sum_i
k_{0,i} = 2$.

The frequency in the initial generation is $x_0 = \nicefrac{1}{N} \sum_{i=1}^N
x_{0,i}$, though to indicate that we treat this as fixed, we use $p_0 = x_0$.
Then, the allele frequency change is,

\begin{align}
  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \\
  \Delta x_1 &= K_1 + H_1
\end{align}

where $K_1$ and $H_1$ are the random change in neutral allele frequency change
due to offspring number (including heritable and non-heritable components), and
Mendelian segregation.

Now, let us look the variance of $\var(\Delta x_1)$ over evolutionary
replicates. Since the Mendelian random process and the offspring random process
are independent,

\begin{align}
  \var(\Delta x_1) &= \var(K_1) + \var(H_1).
\end{align}

Looking at each term,

\begin{align}
  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var\left(x_{0,i} (k_{0,i} - 1) \right)
\end{align}

where we ignore the covariance terms due to the sum, since these are on order
$\nicefrac{1}{N^3}$. In the first generation from an arbitrary starting point,
the neutral alleles assort independently into diploids with respect to their
fitness, so

\begin{align}
  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var(x_{0,i}) \var(k_{0,i} - 1)  \\
            &\approx \frac{1}{4N} \var(x_{0,i}) \var(k_{0,i}).
\end{align}

Assuming that no correlation between parental gametes (e.g. no inbreeding),
$\var(x_{0,i})$ is the binomial variance in individual allele frequency, or
$p_0(1-p_0)/2$, and $\var(k) := \var(k_{0,i})$ is the offspring variance of an
individual given by the reproduction process. For example, if the reproduction
process is a neutral multinomial Wright--Fisher, $\var(k) \approx 2$.

\begin{align}
  \var(K_1) &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4}.
\end{align}

The Mendelian noise variance term can be derived similarly. First, the sum over
each parent's transmitted gametes can be simplified by noting that parents are
exchangeable over evolutionary replicates with respect to their contribution to
this term. In a constant population size, the double summation over parents and
their offspring can be replaced by a summation over offspring, since both sum
$N$ exchangeable terms. Then, note that $\var(\delta_{i,j}) = \nicefrac{1}{4}
p_0(1-p_0)$, so

\begin{align}
  \var(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
\end{align}

Finally, we have

\begin{align}
  \var(\Delta x_1) &= \var(K_1) + \var(H_1) \\
  &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \\
                   &\approx \frac{p_0(1-p_0)}{2N}\left(\frac{\var(k)}{4} + \frac{1}{2}\right)
\end{align}

(c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
Note that if the reproduction process is a neutral multinomial Wright--Fisher,
$\var(k) \approx 2$, and this simplifies to the expected Wright--Fisher
variance. If we were to \emph{define} a variance effective population size by
setting this variance in neutral allele frequency change to the expected
variance under a Wright--Fisher model ($V_\text{WF}$), we'd have

\begin{align}
  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \\
  V_\text{WF} &= \var(\Delta p_1) \\
  N_e &= \frac{4N}{\var(k) + 2} \\
\end{align}

(c.f. \cite{Wright1938-tv}).

Now, we look at this variance equation behaves with $t > 1$.

\begin{align}
  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
\end{align}

and

\begin{align}
  \label{eq:var_wf}
  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + \mathcal{C}_{1,2} \\
          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + \mathcal{C}_{1,2}
\end{align}

Consequently, the variance of the neutral allele frequency changes each
generation according to the probability of failing to coalesce each generation
and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up due to associations
with heritable fitness backgrounds. Note that the covariance term $\mathcal{C}_{1,2} =
\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
evolution, and this simplifies to the well-known equation for variance in a
Wright--Fisher population.

>>>>

Our work is motivated by a well-known limitation of classic background
selection theory: it is only accurate for strongly deleterious mutations
\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}. Given
statistical methods that fit linked selection processes to genome-wide
diversity data rely on the classic BGS model to estimate the distribution of
fitness effects and rate of deleterious mutations, it is important to
characterize and remedy these shortcomings. However, extending background
selection theory to model weak negative selection is challenging, and it is
worthwhile to develop some intuition for why this is the case, how the classic
background selection works, and how the quantitative genetic model of
\textcite{Santiago2016-mu} solves these problems and can be extended to
genome-wide inference.

Both the classic background selection and SC16 models imagine a negative
selection process where a continual influx of deleterious mutations flow into
the population at a rate of $\mu$ per basepair per generation in a conserved
region of $L$ basepairs, such that the region-wide per generation mutation rate
is $U = \mu L$. Each mutation imposes an additive selective cost of $s$ in
heterozygotes and $2s$ in homozygotes. The classic background selection model
only consider mutations that are sufficiently deleterious that they are
destined to extinction
\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1995-pt,Hudson1994-oh}. In
this case, the genealogy is well-approximated by a neutral coalescence process
with a rescaled effective population size of $N_0 = N \exp(-\nicefrac{U}{s})$.
This approximation works reasonably well for two reasons (though see
\cite{Cvijovic2018-vd,Walczak2012-fi,Nicolaisen2012-vs}). First, when selection
is strong, the number of deleterious mutations per haplotype has a stationary
Poisson distribution with an average equal to the equilibrium under the
deterministic mutation-selection model \parencite{Haldane1927-ga}. Since
lineages carrying strongly deleterious mutations cannot contribute to the
genealogy in the long run, coalescence events only occur in the fraction
$\exp(-\nicefrac{U}{s})$ of the population free of mutations. Second, the time
it takes for a present-day lineage to trace their ancestry back to a
mutation-free ancestor (the ``delay phase") is negligible compared to the
coalescence timescale of $N_0$ among these ancestors (the ``coalescence
phase"), and can be ignored (\cite{Durrett2008-ql}, p. 213;
\cite{Good2014-yz}).

When mutations are only weakly deleterious, their allele frequencies are
strongly influenced by stochastic perturbations and their frequency
trajectories are no longer well-approximated by deterministic models. When
genetic drift is the only source of randomness in allele frequency change, the
relevant scale of stochastic perturbations is determined by the drift-effective
population size $N_e$ (e.g. due to only non-selective process). The insight of
nearly neutral theory (XXX) was that the qualitative behavior of selected sites
is determined by the compound parameter $2N_e s$. When $2N_e s \gg 1$, the
stochastic component of allele frequency change is minuscule relative to
selection and can be ignored; this is why the factor that rescales the
drift-effective population size under the BGS model does not depend on $N_e$.
However, when $2N_e s \le 1$, the stochastic perturbations are on a scale equal
to or greater than the changes due to selection. In this regime, weakly
deleterious alleles can drift up to intermediate frequencies before their
eventual loss or fixation. When this occurs, classic BGS theory breaks down for
several reasons. First, the ``delay phase" is no longer negligible as lineages
carrying weakly deleterious mutations can persist on coalescent timescales.
This distorts genealogies away those expected under neutrality
\parencite{Przeworski1999-mb,OFallon2010-my,Higgs1995-xc}. Second, the
distribution of the number of deleterious mutations (and its corresponding
fitness distribution) are no longer stationary and become a traveling wave
\parencite{Rouzine2008-qz,Good2013-lp,Gessler1995-hz} towards reduced
population fitness (since beneficial and back mutations are ignored).
Consequently, individuals in the least-loaded class are stochastically lost,
fixing those mutations with a click of ``Muller's ratchet"
\parencite{Muller1964-ki}. Determining both the rate of Muller's ratchet
\parencite{Haigh1978-gt,Gordo2002-dr,Gessler1995-hz} and the shape of
genealogies under weak selection have been stubborn open problems in
evolutionary genetics.

To further complicate matters, both strong and weakly deleterious mutations
create \emph{heritable} fitness variation. In turn, the presence of heritable
fitness variation generates another source of randomness in allele frequency
change known as genetic \emph{draft} \parencite{Neher2013-dz}. An allele
experiences genetic draft when it becomes randomly associated with a fitness
background, which stochastically perturbs its trajectory across generations.
Like genetic drift, perturbations due to draft are directionless but overall
increase the variance in allele frequency change. Unlike drift, stochastic
perturbations caused by draft are autocorrelated across generations for as long
an allele does not recombine or segregate off its fitness background and the
selective environment does not change
\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When selection is
strong, draft can generate multiple-merger coalescences that create
discontinuities in allele frequency trajectories, making diffusion
approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
However, when selection is weak, draft is well-approximated as an increase in
the rate of drift (or equivalently, a reduction in effective population size).

A key insight from Robertson (\citeyear{Robertson1961-ho}) is that in the
long-run weak draft acts like an increase in the variance in offspring number.
\textcite{Wright1938-tv} showed that additional non-heritable variance in
offspring number (e.g. due to mating strategy) rescales the population size
according to,

@@

While using Equation \eqref{eq:Va} in Equation \eqref{eq:main_Ne} leads to a
prediction for the draft-effective population size $N_d$, closed-form
expressions for the rate of the ratchet $R$ have generally been hard to find
\parencite{Haigh1978-gt,Higgs1995-xc,Gessler1995-hz}. The key insight of
\textcite{Santiago2016-mu} is that the ratchet rate under draft is determined
by the probability of fixation $p_F(N_d, s)$
\parencite{Kimura1962-su,Malecot1948-zv} with the draft-rescaled effective
population size, i.e. $R = N U p_F(N_d, s)$. Given this equation for the
ratchet and Equation \eqref{eq:polygenic_Ne} for $N_d$ under draft, we have a
system of two non-linear equations that can be solved numerically for $N_d$ and
$R$ for each segment,

@@

where $V_a = (U-R)s$ as in Equation \eqref{eq:Va}. The predicted $V_a$, $R$,
and $N_d$ values represent the equilibria expected under a
mutation-selection-drift-draft process for neutral to strongly deleterious
additive mutations.

With the equilibrium $V_a$ calculated at each segment for a particular $\mu$
and $s$, we can estimate the reduction factor $B(x) = \nicefrac{N_d}{N}$ at any
genomic position $x$ (see Methods XXX). XXX

Since it is too computationally-intensive to numerically solve this system of
equations for all segments to estimate the reduction at every position

for every focal position $x$

@@

Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
haploid-model for a single region under equilibrium negative selection, we
first sought to confirm it could accurately predict the reduction in realistic
forward-time simulations of diploid populations with recombination. Using SLiM
\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
basepairs under varying levels of mutation and selection (see Methods
\ref{sec:methods-segsim}). We find a close correspondence between the observed
and predicted reductions in effective population size $B=\nicefrac{N_d}{N}$
over all selection and mutation parameters including weak selection (Figure
\ref{fig:figure-1}A), in contrast to classic BGS theory. Furthermore, to
investigate whether this accuracy was caused by the model correctly predicting
the equilibrium fitness variance and ratchet rate, we also measured these
throughout the simulation. Again, we find the diploid SC16 theory accurately
predicts both the ratchet rate (Figure 1\ref{fig:figure-1}B) and the genic
fitness variance (Figure 1\ref{fig:figure-1}C).

Moreover, these simulations provide intuition about the underlying negative
selection process. When mutations are strongly deleterious, there is no chance
they can fix, and the ratchet rate is zero (Figure \ref{fig:figure-1}B for $2Ns
> 1$). In this strong selection regime, the additive genic fitness variation
closely matches the theoretic deterministic equilibrium of $V_a = Us$ (dashed
gray line, Figure \ref{fig:figure-1}C) However, around $2N_e s \approx 1$, the
ratchet begins clicking as $p_F > 0$. When this occurs, each click of the
ratchet eliminates variation, and the equilibrium variation diverges from the
deterministic mutation-selection equilibrium (Figure \ref{fig:figure-1}C).

@@

Given the accuracy of the SC16 model in predicting the reduction factor $B$ and
ratchet rate for a single segment under general mutation-selection processes,
we extended their model so that it could be applied to whole-genome data. Our
software method \texttt{bgspy} calculates the equilibrium additive genic
fitness variance ($V_a$) and ratchet rate ($R$) across grids of mutation rates
and selection coefficients for each pre-specified segment in the genome that
may be under negative selection (e.g. coding sequences or UTRs). Then, our
method computes a reduction map $B(x)$ for each position $x$ in the genome,
which is determined by the fitness variation at each segment and their spacing
along the user-supplied recombination map (see Methods XXX). To distinguish
between McVicker's B maps based on classic BGS theory, we call our reduction
map $B(x)$ a B' map.

To assess our predicted B' reduction maps, we simulated negative selection for
fixed mutation rates and selection coefficient on human chromosome 10 using
realistic recombination maps and conserved feature positions. Averaging over
$100$ replicates, we estimated the simulation expected reduction map and
compared this to the predicted B and B' maps for the same parameters. We find
that our B' maps and the classic BGS theory B maps closely match simulations
when selection is strong (top row of Figure \ref{fig:figure-1}E). This provides
the first realistic forward-time chromosome-scale simulation confirmation of
classic BGS theory. However, we find slight discrepancies in regions with low
recombination (Figure \ref{fig:figure-1}E). Second, we find our theory is
vastly more accurate than the classic B maps when selection is very weak ($2N_e
s \ll 1$; bottom row of Figure \ref{fig:figure-1}E), which is expected since
BGS theory does not work in this domain. Across all mutation and selection
parameters simulated, the relative error of the classic B maps is 14.6\%
whereas the relative error in the new B' maps is 5\%. Additionally, we find
that the mean squared error between simulations and B' maps is close to the
theoretic lower bound set by the coalescence variance \parencite[see Methods
XXX]{Tajima1983-gu}.

Nearly all of the error between the new B' maps and chromosome-wide simulations
occurs around the drift-barrier domain of $2Ns \approx 1$ (Figure
\ref{fig:figure-1}D). The higher error in this domain occurs across all
mutation rates. We hypothesized that the higher error in this domain may be
because when our method numerically solves Equations \eqref{eq:main_eqns} for
each segment, it does not consider the reduction experienced due to selection
at other linked segments. In particular, we use a fixed drift-effective
$N=1000$ corresponding to the number of diploids in the simulations rather than
$B(x) N$ at position $x$. To test this, we implemented a locally-rescaled
version of the B' maps, which numerically solves Equations \eqref{eq:main_eqns}
for each segment at position $x$ taking into account the reduction experienced
due to selection at other segments by setting $B'(x)N$ as the population size
(see Method XXX). We find the locally-rescaled B' maps reduce the relative
error from 5\% to 0.4\% and mean squared error (Figure \ref{fig:figure-1}D,
dashed colored lines), but does not entirely eliminate the error in the $2Ns
\approx 1$ domain.

@@

While we find evidence that local-recalling reduces error in the $2Ns \approx
1$ domain, we do not employ it in our genome-wide inference method for a few
reasons. First, the reduction map

numerically solving the system of equations

@@

\begin{align}
    \label{eq:sub}
    r_f = \mu \int_0^1 p_F(s) ds
\end{align}

@@

\subsection*{Solving the B' Equations for each Segment}
\label{sec:methods-bprime-eqns}

In our implementation of the theory of \textcite{Santiago2016-mu}, we calculate
the draft-effective population size $N_d = B N$ within a segment by solving a
system of two non-linear equations,

\begin{align}
  \label{eq:}
  {N}_{d} &= N \exp \left( -V_a \frac{Q^2}{2} \right) & \text{\emph{effective population equation}} \\
  R &= \frac{4N_eU s}{\exp(4 N_e s)-1}  & \text{\emph{substitution equation}} 
\end{align}

\begin{align}
  %V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
  V_a &= (U  - 2 R) s & \text{\emph{fitness variance equation}} \\
  Q^2 &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
  Z &= 1 - \frac{Us}{U-{R}} & \text{\emph{variance decay rate}}
\end{align}

and $U = \mu L$ and $M = r_\text{BP} L$ are the total mutation and
recombination rates for a segment $L$ basepairs long.

% Each mutation has a
% selection coefficient of $s$, for which we assume additive effects. Then, the
% local reduction factor is $b(x) = \nicefrac{N_e(x)}{N}$, and the per-basepair
% per-generation substitution rate is $\lambda_D = \nicefrac{R(x)}{L}$. Note that
% in the variance equation, the long-run variance is the balance between mutation
% and twice the loss due to substitution, $(U-2R)s$, since for selective balance
% $V(x) = 0$ for $s>0$, the per-basepair substitution rate must be half the mutation rate,
% since each substitution removes twice the variance introduced by a mutation in
% a diploid species.

Our software \texttt{bgspy} solves these two equations numerically for each
pre-specified segment across grids of mutation rates and selection
coefficients. The recombination rate in a segment is determined by a
user-supplied recombination map; segments that bridge a change in recombination
rate are split, and solved independently with the appropriate rate. This
pre-calculation step is done in parallel, and determines the equilibrium
additive genic fitness variation for each segment under a
mutation-selection-drift-draft process across the grids of parameters.

\subsection*{Calculating the Reduction Maps}
\label{sec:methods-maps}

The pre-computed equilibria are used to calculate the $B(x)$ maps across
parameter grids, which use in our comparisons to chromosome-scale forward
simulations and our composite-likelihood method. The reduction at a focal site
$x$ can be calculated as the product of all segments' contributing reduction
factors accounting for the recombination distances between the segment and
focal site. Additionally, to accommodate the estimation of separate DFEs for
different types of features, each $B(x)$ maps is

\begin{align}
    \log(B(x)) = - \sum_{i=1} \sum_{j=1} 
\end{align}

\subsection*{Segment Simulations}
\label{sec:methods-segsim}

@@

@@
